Follow-up Samples

### 2017-04-19 submission/manuscript.Rmd

The microbiome changes associated with carcinoma were found to be towards those with normal colons.  This would suggest that treatment for CRC is not only successful for removing the carcinoma but also at removing bacteria associated with them.  This is important due to the numerous studies that have pointed towards a link between the community and tumorgenesis [@zackular_manipulation_2016; @zackular_gut_2013; @baxter_structure_2014].  The importance of the community to this process is further highlighted in our data [Figure 2, 4, & S3].  Interestingly, those who had an increase in their positive probability of carcinoma had no change in their median value for oral pathogens found within the carcinoma model but had a decrease in non-oral pathogens [Figure S4].  Although this finding was not statistically significant, this trend would further highlight the importance of community to tumorgenesis.  Future studies with larger power may be able to confirm these reported observations.

Although we were not able to identify specific OTUs due to the heterogeneity of the population.  This is a common problem that occurs in other fields of microbiome study [@sze_looking_2016].  This made it very difficult to identify individual OTUs associated with treatment and recovery towards a normal colon.  To overcome this large individual heterogeneity we used RF models that made it easier to identify consortia of microbiota that were associated with treatment and recovery.  A potential future direction would be to study whether specific gene signatures associated with these consortia of bacteria have a role to play in the recurrence of CRC.

Despite some of these stated shortcomings our findings add to the existing scientific knowledge on CRC and the microbiome: That there is a measurable change in the bacterial community after adenoma, advanced adenoma, or carcinoma treatment. Further, individuals with carcinoma had a larger change than adenoma or advanced adenoma indivdiuals and this change in community was towards those with normal colons. Our data provides additional evidence on the importance of the community  on recovery of the microbiome after CRC treatment and opens interesting avenues of research into how these changes may affect recurrence.

>>>>

The microbiome changes associated with carcinoma were found to be towards those with normal colons.  This would suggest that treatment for CRC is not only successful for removing the carcinoma but also at removing bacteria associated with them.  This is important due to the numerous studies that have pointed towards a link between the community and tumorgenesis [@zackular_manipulation_2016; @zackular_gut_2013; @baxter_structure_2014].  The importance of the community to this process is further highlighted in our data [Figure 2, 4, & S3].  Interestingly, those who had an increase in their positive probability of carcinoma had no change in their median value for oral pathogens within the carcinoma model but had a decrease in bacteria that were higher in carcinoma [Figure S4].  Although this finding was not statistically significant, this trend further highlights the importance of community to tumorgenesis.  Future studies with larger power may be able to confirm these reported observations.

Although we were not able to identify specific OTUs due to the heterogeneity of the population.  This is a common problem that occurs in other fields of microbiome study [@sze_looking_2016].  This makes it very difficult to identify individual OTUs associated with treatment and recovery towards a normal colon.  To overcome this large individual heterogeneity we used RF models that made it easier to identify consortia of microbiota that were associated with treatment and recovery.  One potential future direction would be to study whether specific gene signatures associated with these consortia of bacteria are directly involved with the pathogenesis of CRC recurrence.

Despite some of these stated shortcomings our findings add to the existing scientific knowledge on CRC and the microbiome: That there is a measurable change in the bacterial community after adenoma, advanced adenoma, or carcinoma treatment. Individuals with carcinoma had a larger change than adenoma or advanced adenoma individuals and this change in community was towards those with normal colons. Our data provides additional evidence on the importance of the community  on recovery of the microbiome after CRC treatment and opens interesting avenues of research into how these changes may affect recurrence.
